---
title: "linear models"
author: "Shihui Peng"
date: "2023-11-24"
output: github_document
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## load and clean the Airbnb data

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  select(price, stars, borough = neighbourhood_group, neighbourhood, room_type) |> 
    filter(borough != "Staten Island")
```

# let's fit a model

## fit a model
```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)

# or we can use this:
fit =
  nyc_airbnb |> 
  lm(price ~ stars + borough, data = _)

# let's look at the 'fit'
fit
```
* no need to create dummy variables by ourselves, r will recognize the categorical variables and create for us.
* The `lm` function begins with the formula specification – outcome on the left of the ~ and predictors separated by + on the right. 
  * interactions between variables can be specified using `*`.
  * You can also specify an intercept-only model (`outcome ~ 1`)
  * a model with no intercept (`outcome ~ 0 + ...`) (default is with intercept)
  * a model using all available predictors (`outcome ~ .`).

## we can use these common functions, but usually we don't
```{r}
summary(fit)
summary(fit)$coef
coef(fit)
# fitted.values(fit) --> pull out the fitted values
```
* The reason that we omit the output
  * it’s a huge pain to deal with. `summary` produces an object of class `summary.lm`, which is also a list – that’s how we extracted the coefficients using `summary(fit)$coef`. `coef` produces a vector of coefficient values, and `fitted.values` is a vector of fitted values. None of this is tidy.

## instead, we want tidy up the output (`broom::glance()`) and tidy up the coefficients (`broom::tidy()`):
```{r}
fit |> 
  broom::glance()

fit |> 
  broom::tidy()

# broom::tidy() outputs a table but this outputs a matrix (found based on...)
summary(fit)$coef |> class()
```
* `broom::tidy()` outputs a table vs `summary(fit)$coef` outputs a matrix. so broom::tidy() is better b/c a table is easier to be used.

### why we say a table is better... e.g.:
```{r}
fit |> 
  broom::tidy() |> 
  mutate(
    term = str_replace(term, "^borough", "Borough: ")
  ) |> 
  select(term, estimate, p.value) |> 
  knitr::kable(digits = 3)



```
* a good thing is everything after b::t is familiar tidyverse codes.
* As an aside, broom::tidy works with lots of things, including most of the functions for model fitting you’re likely to run into (survival, mixed models, additive models, …).

### when r creating indicator variables for categorical vars:

```{r}
fit_1 = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)
  ) |> 
  lm(price ~ stars + borough, data = _)

fit_1 |> 
  broom::tidy()
```
* it takes whatever cat var I has started with (e.g. stars before borough here)
* it assumes the same factor order as ggplot does (which is, alphabetical order). but we can mutate it if we need. as below:

### let's fit another model

```{r}
fit = lm(price ~ stars + borough + room_type, data = nyc_airbnb)

fit |> 
  broom::tidy()
```

